Compositional equivalences based on Open pNets

### 2020-02-26 2016-Weak-FH-Bisim/WeakBisim.tex
proofs again

\subsection{Composition properties}\label{sec:app-composition}
This section gives detailed lemmas and their proofs, these are the equivalent of the composition lemmas for open transitions, but applied to weak automata that give the semantics of open pNets.

\begin{lemma}[Weak open transition decomposition]\label{lem-decomposeWOT} 
	Let $\Leaves(Q)=pLTS_l^{l\in L_Q}$; suppose\footnote{Note that the hypotheses of the 
	lemma imply that $Q$ is 
	not a pLTS but a similar lemma can be proven for a pLTS $Q$}:
	\[ P[Q]_{j_0}  
		\models
		{\openrule
			{
				\gamma_j^{j\in J}, \Pred,  
				\Post}
			{\ostate{s_i^{i\in L}} \OTWeakarrow {\alpha}
				\ostate{s_i'^{\, i\in L}}}
		}
	\]
		with  $J\cap\Holes(Q)\neq\emptyset$ or $\exists i\in L_Q.\,s_i\neq s'_i$, i.e. $Q$ takes part in the reduction.  
		 Then there exist $n$, $\Pred\,'$,  
		$\Post\,'$,   and for all $p\in[0..n]$ there exist $\beta_p$, $\alpha_p$, $\Pred_p$, $\Post_p$ and a family $\gamma_{p j}^{j\in J_p}$ and for all $p\in[0..n+1]$ $s_{p i}^{\,i\in L_Q}$. s.t.:\\[-2ex]
		%\[
		\begin{mathpar}
		P\models{\openrule
			{
				\gamma_j^{j\in (J_p\setminus \Holes(Q)) \cup \{j_0\}}, 
				\Pred\,',  
				\Post\,'}
			{\ostate{s_i^{i\in L\setminus L_Q}} \OTWeakarrow {\alpha}
				\ostate{s_i'^{\,i\in L\setminus L_Q}}}
		}%\]
\text{\qquad and \qquad}\gamma_{j_0}=[\beta_0..\beta_n]
	\vspace{-2.2ex}\\\text{and for all $p\in[0..n]$~~}
		%\[
		Q\models{\openrule
			{
				\gamma_{p j}^{j\in J_p}, \Pred_p,  
				\Post_p}
			{\ostate{s_{p i}^{i\in L_Q}} \OTWeakarrow {\alpha_p}
				\ostate{s_{(p+1) i}^{\,i\in L_Q}}}
		}%\]
\\
		\text{such that~\qquad}  \bigcup_{p=0}^nJ_{p} = J\cap\Holes(Q) 
\text{, }
 \gamma_j^{j\in J\cap\Holes(Q)}= \mybigdotcup_{p=0}^n (\gamma_{p j}^{j\in J_p})\subst{\bigotimes_{i=p-1}^{0}\Post_i}, \\
{~\hspace{2cm}}\Pred \iff \Pred\,'
		\land \!\!\bigwedge_{p=0}^n(\alpha_p\subst{\bigotimes_{i=p-1}^{0}\Post_i} = \beta_p\land \Pred_p\subst{\bigotimes_{i=p-1}^{0}\Post_i} ),
\\
{~\hspace{2cm}}\Post=\Post\,' \uplus \bigotimes_{p=n}^0
		\Post_p,   \text{and\quad} \forall i \in L_Q.\, s_{(n+1) i} = s_i'\land s_{0 i} = s_i\\
		\end{mathpar} 
where $\Post_p$ only act upon variables of 
		$\Leaves(Q)$.
\end{lemma}

\begin{proof} Suppose that we have: 
\begin{mathpar}
 P[Q]_{j_0} \models \openrule
			{
				{\gamma_j^{j\in J}}, \Pred,  
				\Post}
			{\ostate{s_l^{l\in L}} \OTWeakarrow {\alpha}
				\ostate{s_l'^{\, l\in L}}} 
\end{mathpar}
By Lemma \ref{lem-rel-OT-WOT} this implies the following: \\			
$ \forall p\! \in [0..m_1]\,		
	P[Q]_{j_0}\!\models		
\openrule
    {
       \set{\beta_{1p}},\Pred_{1p},\Post_{1p}   }
         {\ostate{s^{l\in L}_{pl}} \OTarrow {\tau} \ostate{s^{l\in L}_{(p+1)l}}}$,  $\qquad  P[Q]_{j_0}	\models \openrule
         {
           \set{\beta_2},\Pred_2,\Post_2 }
         {\ostate{t_l^{l\in L}} \OTarrow {\alpha'} \ostate{{t'}_l^{l\in L}}}
$ 
\\
and 
 $\forall p \in [0..m_2] \,\, P[Q]_{j_0}\!\models \openrule
         {
           \set{\beta_{3p}},\Pred_{3p}, \Post_{3p}   }
         {\ostate{u^{l\in L}_{pl}} \OTarrow {\tau} \ostate{u^{l\in L}_{(p+1)l}}}
$\\
where\\
$\alpha=\alpha'\displaystyle{\subst{\bigotimes_{j=m_1}^{0}\!\!\Post_{1j}}}$\\
 for all $l\in L$: $s_l=s_{0 l} \wedge s_{(m_1+1) l}=t_l \wedge t'_l = u_{0 l} \wedge u_{(m_2+1) l}=s'_l$\\
$\displaystyle{
\gamma_j^{j\in J}=\mybigdotcup_{i=0}^{m_1} \vis{\set{\beta_{1i}}\subst{\bigotimes_{j=i-1}^{0}\!\! \Post_{1j}}}\dotcup\  \vis{\set{\beta_2}\subst{\bigotimes_{j=m_1}^{0}\Post_{1j}}}} \dotcup \\ ~\quad\,\qquad \displaystyle{\mybigdotcup_{i=0}^{m_2} \vis{\set{\beta_{3i}} \subst{\bigotimes_{j=i-1}^{0}\!\Post_{3j}\shortotimes\Post_2\shortotimes\bigotimes_{j=m_1}^{0}\!\!\Post_{1j}} }}
\\
\Pred=\bigwedge_{i=0}^{m_1}\Pred_{1i}\subst{\bigotimes_{j=i-1}^{0}\!\!\Post_{1j}}\land\Pred_2 \subst{\bigotimes_{j=m_1}^{0}\!\!\Post_{1j}}\land\\ 
~\qquad\quad\bigwedge_{i=0}^{m_2}\Pred_{3i}\subst{\bigotimes_{j=i-1}^{0}\!\!\Post_{3j}\shortotimes\Post_2\shortotimes\bigotimes_{j=m_1}^{0}\!\!\Post_{1j}}\\
\Post=\bigotimes_{j=m_2}^{0}\!\!\Post_{3j}\shortotimes\Post_2\shortotimes\bigotimes_{j=m_1}^{0}\!\!\Post_{1j}$\\

@@

\end{enumerate}
By using  Lemma \ref{lem-rel-OT-WOT}, and denoting $\displaystyle{J=\bigcup_{i=0}^{m_1}J_{1i}\cup J_2\cup \bigcup_{i=0}^{m_2}J_{3i}}$, we can conclude from cases 1, 2 and 3 that we have:
\begin{mathpar}
		P\models{\openrule
			{
				(\gamma_j')^{j\in (J\setminus \Holes(Q)) \cup \{j_0\}}, 
				\Pred\,',  
				\Post'}
			{\ostate{s_l^{l\in L\setminus L_Q}} \OTWeakarrow {\alpha''}
				\ostate{s_l'^{\,l\in L\setminus L_Q}}}
		}%\]
\end{mathpar}
where:\\
$\alpha''=\alpha'\displaystyle{\subst{\bigotimes_{j=m_1}^{0}\!\!\Post'_{1j}}}$\\
On the other hand, we have:
$\alpha=\alpha'\displaystyle{\subst{\bigotimes_{j=m_1}^{0}\!\!\Post'_{1j}}}\displaystyle{\subst{\bigotimes_{j=m_1}^{0}\!\!\Post''_{1j}}}=\alpha''\displaystyle{\subst{\bigotimes_{j=m_1}^{0}\!\!\Post''_{1j}}}$.\\
As $\displaystyle{\subst{\bigotimes_{j=m_1}^{0}\!\!\Post''_{1j}}}$ has no effect on variables of $P$ and thus on variables of $\alpha$, so we have $\alpha=\alpha''$.
\\
For all $l\in L$: 
$s_l=s_{0l} \wedge s_{(m_1+1)l}=t_l \wedge t'_l = u_{0l} \wedge u_{(m_2+1)l}=s'_l$
and \\%$\forall j\in (J\setminus \Holes(Q))\cup \{j_0\}$,\\
${(\gamma_j')}^{j\in (J\setminus \Holes(Q)) \cup \{j_0\}}=\displaystyle{\mybigdotcup_{i=0}^{m_1} \vis{\set{\beta_{1i}}\subst{\bigotimes_{j=i-1}^{0}\!\!\Post\,'_{1j}}}\dotcup  \vis{\set{\beta_{2}}\subst{\bigotimes_{j=m_1}^{0}\Post\,'_{1j}}}} \dotcup\\
~~\qquad \qquad\qquad\qquad\qquad \mybigdotcup_{i=0}^{m_2}
\vis{\set{\beta_{3i}} \subst{\bigotimes_{j=i-1}^{0}\!\Post\,'_{3j}\shortotimes\Post\,'_2\shortotimes\bigotimes_{j=m_1}^{0}\!\!\Post\,'_{1j}} }\\
\Pred\,'=\bigwedge_{i=0}^{m_1}\Pred\,'_{1i}\subst{\bigotimes_{j=i-1}^{0}\!\!\Post\,'_{1j}}\land\Pred\,'_2 \subst{\bigotimes_{j=m_1}^{0}\!\!\Post\,'_{1j}}\land\\ 
~\qquad\qquad\bigwedge_{i=0}^{m_2}\Pred\,'_{3i}\subst{\bigotimes_{j=i-1}^{0}\!\!\Post\,'_{3j}\shortotimes\Post\,'_2\shortotimes\bigotimes_{j=m_1}^{0}\!\!\Post\,'_{1j}}\\
\Post'=\bigotimes_{j=m_2}^{0}\!\!\Post\,'_{3j}\shortotimes\Post\,'_2\shortotimes\bigotimes_{j=m_1}^{0}\!\!\Post\,'_{1j}$

Note that for all $j\in J\setminus \Holes(Q)$, $\gamma_j' = \gamma_j$ because for all $l$ $\Post'_{1l}$ coincide with $\Post_{1l}$ on the variables of $\beta_{1ij}$, and similarly for $\Post'_2$ and $\Post'_{3l}$.
%Concerning $\gamma'_{j_0}$ we have:
%\[\gamma'_{j_0}=\displaystyle{\mybigdotcup_{i=0}^{m_1} \vis{{\alpha_{1i}}\subst{\bigotimes_{j=i-1}^{0}\!\!\Post'_{1j}}}\dotcup  \vis{\set{\beta_{2}}\subst{\bigotimes_{j=m_1}^{0}\Post'_{1j}}}} \dotcup\\
%~~\qquad \mybigdotcup_{i=0}^{m_2}
%\vis{\set{\beta_{3i}} \subst{\bigotimes_{j=i-1}^{0}\!\Post'_{3j}\shortotimes\Post'_2\shortotimes\bigotimes_{j=m_1}^{0}\!\!\Post'_{1j}} }\]

We denote the following predicate (we will need it for reasoning on the global predicate and will reason on it along the proof):\\
$\begin{array}{l}\displaystyle{
\Pred_\beta=\bigwedge_{p=0}^{m_1}(\beta_{1pj_0}=\alpha_{1p})\subst{\bigotimes_{j=p-1}^{0}\!\!\Post_{1j}} \wedge (\beta_{2j_0}=\alpha_2)\subst{\bigotimes_{j=m_1}^{0}\!\!\Post_{1j}}} \wedge \\\qquad\qquad
\displaystyle{
\bigwedge_{p=0}^{m_2}(\beta_{3pj_0}=\alpha_{3p})
\subst{\bigotimes_{j=p-1}^{0}\!\!\Post_{3j}\shortotimes\Post_2\shortotimes\bigotimes_{j=m_1}^{0}\!\!\Post_{1j}}}
\end{array}$

>>>>

\subsection{Composition properties}\label{sec:app-composition}
This section gives decomposition/composition lemmas and their proofs, these are the equivalent of the composition lemmas for open transitions, but applied to weak open automata.
\begin{lemma}[Weak open transition decomposition]\label{lem-decomposeWOT} 
	Let $\Leaves(Q)=pLTS_l^{l\in L_Q}$; suppose\footnote{Note that the hypotheses of the 
	lemma imply that $Q$ is 
	not a pLTS but a similar lemma can be proven for a pLTS $Q$}:
	\[ P[Q]_{j_0}  
		\models
		{\openrule
			{
				\gamma_j^{j\in J}, \Pred,  
				\Post}
			{\ostate{s_i^{i\in L}} \OTWeakarrow {\alpha}
				\ostate{s_i'^{\, i\in L}}}
		}
	\]
		with  $J\cap\Holes(Q)\neq\emptyset$ or $\exists i\in L_Q.\,s_i\neq s'_i$, i.e. $Q$ takes part in the reduction.  
		 Then there exist $n$, $\Pred\,'$,  
		$\Post\,'$,   and for all $p\in[0..n]$ there exist $\beta_p$, $\alpha_p$, $\Pred_p$, $\Post_p$ and a family $\gamma_{p j}^{j\in J_p}$ and for all $p\in[0..n+1]$ $s_{p i}^{\,i\in L_Q}$. s.t.:\\[-2ex]
		%\[
		\begin{mathpar}
		P\models{\openrule
			{
				\gamma_j^{j\in (J_p\setminus \Holes(Q)) \cup \{j_0\}}, 
				\Pred\,',  
				\Post\,'}
			{\ostate{s_i^{i\in L\setminus L_Q}} \OTWeakarrow {\alpha}
				\ostate{s_i'^{\,i\in L\setminus L_Q}}}
		}%\]
\text{\qquad and \qquad}\gamma_{j_0}=[\beta_0..\beta_n]
	\vspace{-2.2ex}\\\text{and for all $p\in[0..n]$~~}
		%\[
		Q\models{\openrule
			{
				\gamma_{p j}^{j\in J_p}, \Pred_p,  
				\Post_p}
			{\ostate{s_{p i}^{i\in L_Q}} \OTWeakarrow {\alpha_p}
				\ostate{s_{(p+1) i}^{\,i\in L_Q}}}
		}%\]
\\
		\text{such that~\qquad}  \bigcup_{p=0}^nJ_{p} = J\cap\Holes(Q) 
\text{, }
 \gamma_j^{j\in J\cap\Holes(Q)}= \mybigdotcup_{p=0}^n (\gamma_{p j}^{j\in J_p})\subst{\bigotimes_{i=p-1}^{0}\Post_i}, \\
{~\hspace{2cm}}\Pred \iff \Pred\,'
		\land \!\!\bigwedge_{p=0}^n(\alpha_p\subst{\bigotimes_{i=p-1}^{0}\Post_i} = \beta_p\land \Pred_p\subst{\bigotimes_{i=p-1}^{0}\Post_i} ),
\\
{~\hspace{2cm}}\Post=\Post\,' \uplus \bigotimes_{p=n}^0
		\Post_p,   \text{and\quad} \forall i \in L_Q.\, s_{(n+1) i} = s_i'\land s_{0 i} = s_i\\
		\end{mathpar} 
where $\Post_p$ only act upon variables of 
		$\Leaves(Q)$.
\end{lemma}

\begin{proof} Suppose that we have: 
\begin{mathpar}
 P[Q]_{j_0} \models \openrule
			{
				{\gamma_j^{j\in J}}, \Pred,  
				\Post}
			{\ostate{s_l^{l\in L}} \OTWeakarrow {\alpha}
				\ostate{s_l'^{\, l\in L}}} 
\end{mathpar}
By Lemma \ref{lem-rel-OT-WOT} this implies the following: \\			
$ \forall p\! \in [0..m_1]\,		
	P[Q]_{j_0}\!\models		
\openrule
    {
       \set{\beta_{1p}},\Pred_{1p},\Post_{1p}   }
         {\ostate{s^{l\in L}_{pl}} \OTarrow {\tau} \ostate{s^{l\in L}_{(p+1)l}}}$,  $\qquad  P[Q]_{j_0}	\models \openrule
         {
           \set{\beta_2},\Pred_2,\Post_2 }
         {\ostate{t_l^{l\in L}} \OTarrow {\alpha'} \ostate{{t'}_l^{l\in L}}}
$ 
\\
and 
 $\forall p \in [0..m_2] \,\, P[Q]_{j_0}\!\models \openrule
         {
           \set{\beta_{3p}},\Pred_{3p}, \Post_{3p}   }
         {\ostate{u^{l\in L}_{pl}} \OTarrow {\tau} \ostate{u^{l\in L}_{(p+1)l}}}
$\\
where\\
 for all $l\in L$: $s_l=s_{0 l} \wedge s_{(m_1+1) l}=t_l \wedge t'_l = u_{0 l} \wedge u_{(m_2+1) l}=s'_l$\\
$\alpha=\alpha'\displaystyle{\subst{\bigotimes_{j=m_1}^{0}\!\!\Post_{1j}}}$\\
$\displaystyle{
\gamma_j^{j\in J}=\mybigdotcup_{i=0}^{m_1} \vis{\set{\beta_{1i}}\subst{\bigotimes_{j=i-1}^{0}\!\! \Post_{1j}}}\dotcup\  \vis{\set{\beta_2}\subst{\bigotimes_{j=m_1}^{0}\Post_{1j}}}} \dotcup \\ 
{}~\quad\,\qquad \displaystyle{\mybigdotcup_{i=0}^{m_2} \vis{\set{\beta_{3i}} \subst{\bigotimes_{j=i-1}^{0}\!\Post_{3j}\shortotimes\Post_2\shortotimes\bigotimes_{j=m_1}^{0}\!\!\Post_{1j}} }}
\\
\Pred=\bigwedge_{i=0}^{m_1}\Pred_{1i}\subst{\bigotimes_{j=i-1}^{0}\!\!\Post_{1j}}\land\Pred_2 \subst{\bigotimes_{j=m_1}^{0}\!\!\Post_{1j}}\land\\ 
{}~\qquad\quad\bigwedge_{i=0}^{m_2}\Pred_{3i}\subst{\bigotimes_{j=i-1}^{0}\!\!\Post_{3j}\shortotimes\Post_2\shortotimes\bigotimes_{j=m_1}^{0}\!\!\Post_{1j}}\\
\Post=\bigotimes_{j=m_2}^{0}\!\!\Post_{3j}\shortotimes\Post_2\shortotimes\bigotimes_{j=m_1}^{0}\!\!\Post_{1j}$\\

@@

\end{enumerate}
By using  Lemma \ref{lem-rel-OT-WOT}, and denoting $\displaystyle{J=\bigcup_{i=0}^{m_1}J_{1i}\cup J_2\cup \bigcup_{i=0}^{m_2}J_{3i}}$, we can conclude from cases 1, 2 and 3 that we have:
\begin{mathpar}
		P\models{\openrule
			{
				(\gamma_j')^{j\in (J\setminus \Holes(Q)) \cup \{j_0\}}, 
				\Pred\,',  
				\Post'}
			{\ostate{s_l^{l\in L\setminus L_Q}} \OTWeakarrow {\alpha''}
				\ostate{s_l'^{\,l\in L\setminus L_Q}}}
		}%\]
\end{mathpar}
where
$\alpha''=\alpha'\displaystyle{\subst{\bigotimes_{j=m_1}^{0}\!\!\Post\,'_{1j}}}$\\
On the other hand, we have:
$\alpha=\alpha'\displaystyle{\subst{\bigotimes_{j=m_1}^{0}\!\!\Post\,'_{1j}}}\displaystyle{\subst{\bigotimes_{j=m_1}^{0}\!\!\Post\,''_{1j}}}=\alpha''\displaystyle{\subst{\bigotimes_{j=m_1}^{0}\!\!\Post\,''_{1j}}}$.\\
As $\displaystyle{\subst{\bigotimes_{j=m_1}^{0}\!\!\Post\,''_{1j}}}$ has no effect on variables of $P$ and thus on variables of $\alpha$, so we have $\alpha=\alpha''$.
\\
For all $l\in L$: 
$s_l=s_{0l} \wedge s_{(m_1+1)l}=t_l \wedge t'_l = u_{0l} \wedge u_{(m_2+1)l}=s'_l$
and \\%$\forall j\in (J\setminus \Holes(Q))\cup \{j_0\}$,\\
${(\gamma_j')}^{j\in (J\setminus \Holes(Q)) \cup \{j_0\}}=\displaystyle{\mybigdotcup_{i=0}^{m_1} \vis{\set{\beta_{1i}}\subst{\bigotimes_{j=i-1}^{0}\!\!\Post\,'_{1j}}}\dotcup  \vis{\set{\beta_{2}}\subst{\bigotimes_{j=m_1}^{0}\Post\,'_{1j}}}} \dotcup\\
{}~~\qquad \qquad\qquad\qquad\qquad \mybigdotcup_{i=0}^{m_2}
\vis{\set{\beta_{3i}} \subst{\bigotimes_{j=i-1}^{0}\!\Post\,'_{3j}\shortotimes\Post\,'_2\shortotimes\bigotimes_{j=m_1}^{0}\!\!\Post\,'_{1j}} }\\
\Pred\,'=\bigwedge_{i=0}^{m_1}\Pred\,'_{1i}\subst{\bigotimes_{j=i-1}^{0}\!\!\Post\,'_{1j}}\land\Pred\,'_2 \subst{\bigotimes_{j=m_1}^{0}\!\!\Post\,'_{1j}}\land\\ 
{}~\qquad\qquad\bigwedge_{i=0}^{m_2}\Pred\,'_{3i}\subst{\bigotimes_{j=i-1}^{0}\!\!\Post\,'_{3j}\shortotimes\Post\,'_2\shortotimes\bigotimes_{j=m_1}^{0}\!\!\Post\,'_{1j}}\\
\Post'=\bigotimes_{j=m_2}^{0}\!\!\Post\,'_{3j}\shortotimes\Post\,'_2\shortotimes\bigotimes_{j=m_1}^{0}\!\!\Post\,'_{1j}$

Note that for all $j\in J\setminus \Holes(Q)$, $\gamma_j' = \gamma_j$ because for all $l$ $\Post\,'_{1l}$ coincide with $\Post_{1l}$ on the variables of $\beta_{1ij}$, and similarly for $\Post\,'_2$ and $\Post\,'_{3l}$.
%Concerning $\gamma'_{j_0}$ we have:
%\[\gamma'_{j_0}=\displaystyle{\mybigdotcup_{i=0}^{m_1} \vis{{\alpha_{1i}}\subst{\bigotimes_{j=i-1}^{0}\!\!\Post'_{1j}}}\dotcup  \vis{\set{\beta_{2}}\subst{\bigotimes_{j=m_1}^{0}\Post'_{1j}}}} \dotcup\\
%~~\qquad \mybigdotcup_{i=0}^{m_2}
%\vis{\set{\beta_{3i}} \subst{\bigotimes_{j=i-1}^{0}\!\Post'_{3j}\shortotimes\Post'_2\shortotimes\bigotimes_{j=m_1}^{0}\!\!\Post'_{1j}} }\]

We introduce the following predicate (we will need it for reasoning on the global predicate and will reason on it along the proof):\\
$\begin{array}{l}\displaystyle{
\Pred_\beta=\bigwedge_{p=0}^{m_1}(\beta_{1pj_0}=\alpha_{1p})\subst{\bigotimes_{j=p-1}^{0}\!\!\Post_{1j}} \wedge (\beta_{2j_0}=\alpha_2)\subst{\bigotimes_{j=m_1}^{0}\!\!\Post_{1j}}} \wedge \\\qquad\qquad
\displaystyle{
\bigwedge_{p=0}^{m_2}(\beta_{3pj_0}=\alpha_{3p})
\subst{\bigotimes_{j=p-1}^{0}\!\!\Post_{3j}\shortotimes\Post_2\shortotimes\bigotimes_{j=m_1}^{0}\!\!\Post_{1j}}}
\end{array}$
